---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# EBMAhelper

EBMAhelper is a wrapper around EBMAforecast that uses conventional R model syntax and methods, i.e. implementing `ebma()` to create and EBMA ensemble and a standalong `predict()` method to aggregate new forecasts. 

## Example

A EBMA ensemble can be fitted like this:

```{r example}
suppressMessages({
  library("EBMAforecast")
  library("EBMAhelper")
})


data("presidentialForecast")

head(presidentialForecast)
str(presidentialForecast)

fitted_ensemble <- ebma(y = presidentialForecast[ ,7],
                        x = presidentialForecast[ ,c(1:6)],
                        model_type = "normal")
summary(fitted_ensemble)
```

To do the same with `EBMAforecast` is slightly more verbose and does not follow R's model fitting conventions. The following code is adapted from the package demo:

```{r}
data("presidentialForecast")

full.forecasts<-presidentialForecast[,c(1:6)]
full.observed<-presidentialForecast[,7]

this.ForecastData <- makeForecastData(
  .predCalibration=full.forecasts[1:12,],
  .outcomeCalibration=full.observed[1:12],
  .predTest=full.forecasts[13:15,], 
  .outcomeTest=full.observed[13:15], 
  .modelNames=c("Campbell", "Lewis-Beck","EWT2C2","Fair","Hibbs","Abramowitz"))
thisEnsemble<-calibrateEnsemble(this.ForecastData, model="normal", 
                                useModelParams=FALSE, tol = 0.000000001,
                                const = 0)
summary(thisEnsemble)
```

There is also no standalone method for predicting new cases with a fitted EBMA model in `EBMAforecast`. 

## Installation

```{r installation, eval = FALSE}
library("devtools")

install_github("andybega/EBMAhelper")
```

