---
title: "EBMAhelper"
author: "Andreas Beger"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

EBMAhelper is a small helper package that wraps EBMAforecast's ensemble-fitting syntax in more R-conventional model fitting syntax. The function `ebma()` will calibrate an EBMA ensemble model, and the resulting oject can then be used with conventional `predict()`, `print()`, and `summary()` methods, unlike in EBMAforecast. 

## Demonstrations with EBMAforecast demo data

For each of the two included demo datasets, we will fit (calibrate) an ensemble model and predict on a set of holdout test data afterwards. 

```{r}
suppressPackageStartupMessages({
  library("EBMAforecast")
  library("EBMAhelper")
  library("dplyr")
})
```


### Normal model: presidential forecasts

```{r}
data(presidentialForecast)
head(presidentialForecast) %>%
  knitr::kable(digits = 3)
```

The data consists of 6 different sets of forecasts for 15 US presidential elections, as well as teh observed (actual) outcome. Next we split out the input forecasts and observed outcome, and hold the last 3 presidential elections back for out-of-sample test.

```{r}
input_forecasts <- presidentialForecast[, c(1:6)]
outcome <- presidentialForecast[, 7]
train_idx <- 1:12
test_idx  <- 13:15
model_names <- c("Campbell", "Lewis-Beck","EWT2C2","Fair","Hibbs","Abramowitz")
```

Fitting a model with EBMAforecast, adapted from the package demo:

```{r}
this.ForecastData <- makeForecastData(
  .predCalibration    = input_forecasts[train_idx, ],
  .outcomeCalibration = outcome[train_idx],
  .predTest    = input_forecasts[test_idx, ], 
  .outcomeTest = outcome[test_idx], 
  .modelNames = model_names)

ebma_fit <- calibrateEnsemble(this.ForecastData, model="normal")
ebma_fit
summary(ebma_fit, period = "test")
```

The test predictions are calculated as part of `calibrateEnsemble`. If we wanted to predict on a different set of data, we can use `EBMApredict` (see `?EBMApredict`):

```{r}
preds <- EBMApredict(ebma_fit, as.matrix(input_forecasts[test_idx, ]), 
                     Outcome = outcome[test_idx])
class(preds)
preds
```

The predictions come in a S4 object with class "FDatFitNormal"; the actual predictions for EBMA  are in an array in slot "predTest":

```{r}
class(preds@predTest)
# get the EBMA predictions
preds@predTest[, "EBMA", 1]
```

Now the equivalent sequence of steps with EBMAhelper. 

```{r}
ebma_fit_helper <- ebma(y = outcome[train_idx], x = input_forecasts[train_idx, ],
                        model_type = "normal")
class(ebma_fit_helper)
```

The class is different, but this is just a shallow cover; print and summary still work the same. By default `ebma()` will take the "x" column names as model names.

```{r}
summary(ebma_fit_helper)
```

Get EBMA predictions for the test period, and make sure they match:

```{r}
preds_helper <- predict(ebma_fit_helper, newdata = input_forecasts[test_idx, ])
cbind(
  EBMAforecast = preds@predTest[, "EBMA", 1],
  EBMAhelper = preds_helper
)
```

If "newdata" is left at the default NULL value, it will return in-sample predictions for the calibration period.

```{r}
cbind(
  EBMAforecast = ebma_fit@predCalibration[, "EBMA", 1],
  EBMAhelper = predict(ebma_fit_helper)
)
```


### Logit model: insurgency forecasts

The data consist of 3 streams of forecasts for a binary yes/no insurgency indicator. There are 696 rows in total. 

```{r}
data(calibrationSample)
calibrationSample <- as.data.frame(calibrationSample)
head(calibrationSample) %>%
  knitr::kable(digits = 4)
```

Keep the last 20% of rows as holdout test sample; split data accordingly.

```{r}
train_idx <- 1:(ceiling(nrow(calibrationSample)*.8))
test_idx  <- max(train_idx):nrow(calibrationSample)
y_col <- 4
x_col <- 1:3
train_data <- calibrationSample[train_idx, ]
test_data  <- calibrationSample[test_idx, ]
model_names <- c("LMER", "SAE", "GLM")
```


```{r}
this.ForecastData <- makeForecastData(
  .predCalibration    = train_data[, x_col],
  .outcomeCalibration = train_data[, y_col],
  .predTest    = test_data[, x_col],
  .outcomeTest = test_data[, y_col],
  .modelNames = model_names)

ebma_fit <- calibrateEnsemble(this.ForecastData, model="logit")
summary(ebma_fit)
```

This errors out right now:

```{r, error=TRUE}
preds <- EBMApredict(ebma_fit, as.matrix(test_data[, x_col]), 
                     Outcome = test_data[, y_col])
preds <- preds@predTest[, "EBMA", 1]
```

Now with EBMAhelper:

```{r}
ebma_fit_helper <- ebma(y = train_data[, y_col],
                        x = train_data[, x_col],
                        model_type = "logit")
summary(ebma_fit_helper)
```

Skip the next part until predict works for logit.

```{r, eval=FALSE}
preds_helper <- predict(ebma_fit_helper)
cbind(
  EBMAforecast = preds,
  EBMAhelper = preds_helper
) %>%
  head() %>%
  knitr::kable(digits = 3)
```

